---
title: "Diamond Prices I"
author: "Jonathan Freeland"
date: "March 25, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
baseFill <- "#69D2E7"
baseOutline <- "#666666"
boxPlotFill <- "#A7DBD8"
```

Load the diamonds data set.
```{r, cache=TRUE}
data(diamonds)
```

Analyze diamond prices.
```{r, cache=TRUE}
summary(diamonds$price)
diamonds %>% ggplot(aes(x = price)) +
  geom_histogram(binwidth = 250, color = baseOutline, fill = baseFill) +
  scale_x_continuous(breaks=seq(300,19000,2000)) +
  ggtitle("Prices")
```

* There are `r diamonds %>% filter(price < 500) %>% count()` diamonds that cost < $500
* There are `r diamonds %>% filter(price < 250) %>% count()` diamonds that cost < $250
* There are `r diamonds %>% filter(price >= 15000) %>% count()` diamonds that cost >= $15,000

Drill down to visualize the largest price peak.
```{r, cache=TRUE}
diamonds %>% ggplot(aes(x = price)) +
  geom_histogram(binwidth = 10, color = baseOutline, fill = baseFill, na.rm = TRUE) +
  scale_x_continuous(lim=c(300,1500), breaks=seq(300,1500,100)) +
  ggtitle("Price Detail, $300-$1500")
```

Investigate diamond prices by cut.
```{r, cache=TRUE}
diamonds %>% ggplot(aes(x = price)) +
  geom_histogram(binwidth = 500, color = baseOutline, fill = baseFill) +
  facet_wrap(~cut, scales="free_y") +
  ggtitle("Price by Cut")
```

Investigate diamond prices by price per carat and cut.
```{r, cache=TRUE}
diamonds %>% ggplot(aes(x = price/carat)) +
  geom_histogram(binwidth = .05, color = baseOutline, fill = baseFill) +
  scale_x_log10() +
  facet_wrap(~cut, scales="free_y") +
  ggtitle("Price/Carat by Cut")
```

Investigate diamond prices using some boxplots and numerical summaries.
```{r, cache=TRUE}
by(diamonds$price, diamonds$color, summary, digits = max(getOption('digits')))
```
The IQR for diamonds with worst color J is `r IQR(filter(diamonds, color == "J")$price)`, with best color D is `r IQR(filter(diamonds, color == "D")$price)`.
```{r, cache=TRUE}
diamonds %>% ggplot(aes(y = price, x = cut)) +
 geom_boxplot(color = baseOutline, fill = boxPlotFill, outlier.colour = "#F38630") + 
 scale_y_log10() +
  ggtitle("Price by Cut")
diamonds %>% ggplot(aes(y = price, x = clarity)) +
 geom_boxplot(color = baseOutline, fill = boxPlotFill, outlier.colour = "#F38630") + 
 scale_y_log10() +
  ggtitle("Price by Clarity")
diamonds %>% ggplot(aes(y = price, x = color)) +
  geom_boxplot(color = baseOutline, fill = boxPlotFill, outlier.colour = "#F38630") + 
  scale_y_log10() +
  ggtitle("Price by Color")
```

```{r, cache=TRUE}
by(diamonds$price/diamonds$carat, diamonds$color, summary, digits = max(getOption('digits')))
diamonds %>% ggplot(aes(y = price/carat, x = color)) +
  geom_boxplot(color = baseOutline, fill = boxPlotFill, outlier.colour = "#F38630") + 
  scale_y_log10() +
  ggtitle("Price/Carat by Color")
```

Investigate the weight of diamonds.
```{r, cache=TRUE}
diamonds %>% ggplot(aes(x = carat, y = ..count..)) + 
  geom_freqpoly(binwidth = 0.01) +
  scale_x_continuous(breaks=seq(0,5,.25))
```

